A Study on Common Bug Fix Patterns in Rust

### 2022-10-16 main.tex
added abstract/conclusion; fixed results a bit

\begin{abstract}
Insert your abstract here. Include keywords, PACS and mathematical
subject classification numbers as needed.
\keywords{First keyword \and Second keyword \and More}
% \PACS{PACS code1 \and PACS code2 \and more}
\subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

>>>>

\begin{abstract}
    Rust is a relatively new programming language that allows programmers to write programs that have low-level control over resources, as well as high-level safety guarantees. This dual nature is made possible through the ownership framework which has been incorporated in the language. The main component governing over this framework is called Borrow-Checker. Rust has been ranked first for the sixth consecutive year in the Stack Overflow's annual Developer Survey as the most loved programming language. The number of Rust developers is growing as the need for faster and safer software increases. However, to the best of our knowledge, no research has been developed to automatically find the most pervasive bug fix patterns within Rust programs. In this project, using a novel program modelling method, we mined the top 18 most starred Rust projects in Github to discover the most common bug fix patterns lied within their code changes. We stored 87,726 code changes drawn from 57,214 commits across these 18 projects. After clustering the code changes, and conducting a manual analysis, we discovered 19 groups of cross-project bug fix patterns. We introduce these patterns in two groups: (1) general patterns and (2) borrow-checker related patterns.
\keywords{First keyword \and Second keyword \and More}
% \PACS{PACS code1 \and PACS code2 \and more}
\subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

### 2022-10-16 sections/introduction.tex
added abstract/conclusion; fixed results a bit

Producing bugs has been an unavoidable part of computer programming. Researchers have been trying to help software engineers to be more aware of the bug producing practices in development and how to step towards creating bug-free software. One useful step towards that goal would be studying common bug patterns (and their corresponding fixes) that occur across multiple software systems. Research efforts for categorizing common bug patterns date back to 1975 where Endres tried to categorize various bug patterns which occurred in operating system implementations \cite{endres1975analysis}.

In this work, we empirically analyze the most popular Rust projects in GitHub, aiming to discover common bug patterns that these projects share. Our interest comes from a desire to develop program repair tools, but we believe these insights can also be of broader service. Automated program repair \cite{goues2019automated} \cite{liu2018survey} tools try to locate bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit). A powerful automated program repair system has many useful applications and could significantly simplify debugging, thus reducing software development cost \cite{le2012systematic}. A set of common bug fix patterns can be a useful input to an APR system: it directly provides information about fixing similar buggy structures, which can significantly reduce the search space for patch generation \cite{jeffrey2009bugfix}.

Similarly, if we are observing undesired behaviour from our program, that would indicate that there lies a fault within our program. But where is the fault? Fault localization \cite{wong2016survey} is the process of automatically finding the faulty statement(s) in our program. If a fault localization tool knows about the common bug patterns that can appear in the underlying programming language, it can prioritize code segment candidates based on their potentials of producing bugs. That would result in a more robust tool that leverages historical data.

Moreover, knowing which code patterns in a programming language are more susceptible to bugs can help programmers to be more cautious while working with similar code structures. This knowledge can make the developers avoid certain mistakes during production which ultimately would result in reducing the frequent bug patterns within the program. As Monperrus describes, the fewer frequent bug patterns we keep in our software, the less cost we would have to pay for software maintenance \cite{monperrus2014critical}.

Moving further afield, researchers have been exploiting deep learning techniques to learn about the semantics of computer programs and create program modifier models that have a specific goal (e.g. bug fix) \cite{alon2019code2vec} \cite{alon2018general} \cite{raychev2016learning} \cite{bielik2016phog}. However, most learning tools get fixed-size vectors as input and map them to outputs. Unlike image data, it is challenging to compress programs into vectors without any semantic information loss. Current approaches try to define the vector dimensions in a way that accounts for all possible abstract syntax tree forms. That results in a large, sparse space for the inputs. Assuming that the deep learning model we are trying to develop is associated with bug fixing tasks, then knowledge about common bug fix patterns can help prune the input space towards containing fewer dimensions. That can help with creating more efficient models.

Many studies have investigated detecting and categorizing bug fix patterns in general \cite{islam2021changes} \cite{madeiral2018towards} \cite{pan2009toward}, or for a specific programming language \cite{yang2022mining} \cite{hanam2016discovering} \cite{campos2019discovering}. In this work, we study common bug patterns that appear in Rust projects. Rust is a relatively new language that allows programmers to write programs that have low-level control over resources, as well as high-level safety guarantees \cite{klabnik2019rust}. What distinguishes Rust from the other systems programming languages, e.g. C/C++, is its ownership framework, which is an integral part of Rust. The ownership framework keeps track of references to different memory locations, and will free them once they go out of scope. If a program passes the compiler’s checks, it must meet several safety properties; e.g. it has no dangling references or double-frees. These problems had to be manually ensured in languages like C/C++. Moreover, not only does the ownership concept remove the need for a garbage collector in the runtime environment, it does not require the programmer to explicitly do certain safety tasks \cite{qin2020understanding}. Due to the memory and thread safety guarantees that Rust provides, many software companies adopted Rust in production. For example, Dropbox decided to migrate to Rust for developing their storage system, since they were not completely satisfied with the features provided by AWS’s S3 or Golang \cite{dropbox}.

>>>>

Producing bugs has been an unavoidable part of computer programming. Researchers have been trying to help software engineers to be more aware of the bug producing practices in development and how to step towards creating bug-free software. One useful step towards that goal would be studying common bug patterns (and their corresponding fixes) that occur across multiple software systems. Research efforts for categorizing common bug patterns date back to 1975 where Endres tried to categorize various bug patterns which occurred in operating system implementations~\cite{endres1975analysis}.

In this work, we empirically analyze the most popular Rust projects in GitHub, aiming to discover common bug patterns that these projects share. Our interest comes from a desire to develop program repair tools, but we believe these insights can also be of broader service. Automated program repair~\cite{goues2019automated,liu2018survey} tools try to locate bug locations in the source code (using a fault localization module) and then generate patches to fix them, possibly without human intervention, so that the modified program meets desired specifications (implicit or explicit). A powerful automated program repair system has many useful applications and could significantly simplify debugging, thus reducing software development cost~\cite{le2012systematic}. A set of common bug fix patterns can be a useful input to an APR system: it directly provides information about fixing similar buggy structures, which can significantly reduce the search space for patch generation~\cite{jeffrey2009bugfix}.

Similarly, if we are observing undesired behaviour from our program, that would indicate that there lies a fault within our program. But where is the fault? Fault localization~\cite{wong2016survey} is the process of automatically finding the faulty statement(s) in our program. If a fault localization tool knows about the common bug patterns that can appear in the underlying programming language, it can prioritize code segment candidates based on their potentials of producing bugs. That would result in a more robust tool that leverages historical data.

Moreover, knowing which code patterns in a programming language are more susceptible to bugs can help programmers to be more cautious while working with similar code structures. This knowledge can make the developers avoid certain mistakes during production which ultimately would result in reducing the frequent bug patterns within the program. As Monperrus describes, the fewer frequent bug patterns we keep in our software, the less cost we would have to pay for software maintenance~\cite{monperrus2014critical}.

Moving further afield, researchers have been exploiting deep learning techniques to learn about the semantics of computer programs and create program modifier models that have a specific goal (e.g. bug fix)~\cite{alon2019code2vec,alon2018general,raychev2016learning,bielik2016phog}. However, most learning tools get fixed-size vectors as input and map them to outputs. Unlike image data, it is challenging to compress programs into vectors without any semantic information loss. Current approaches try to define the vector dimensions in a way that accounts for all possible abstract syntax tree forms. That results in a large, sparse space for the inputs. Assuming that the deep learning model we are trying to develop is associated with bug fixing tasks, then knowledge about common bug fix patterns can help prune the input space towards containing fewer dimensions. That can help with creating more efficient models.

Many studies have investigated detecting and categorizing bug fix patterns in general~\cite{islam2021changes,madeiral2018towards,pan2009toward}, or for a specific programming language~\cite{yang2022mining,hanam2016discovering,campos2019discovering}. In this work, we study common bug patterns that appear in Rust projects. Rust is a relatively new language that allows programmers to write programs that have low-level control over resources, as well as high-level safety guarantees~\cite{klabnik2019rust}. What distinguishes Rust from the other systems programming languages, e.g. C/C++, is its ownership framework, which is an integral part of Rust. The ownership framework keeps track of references to different memory locations, and will free them once they go out of scope. If a program passes the compiler’s checks, it must meet several safety properties; e.g. it has no dangling references or double-frees. These problems had to be manually ensured in languages like C/C++. Moreover, not only does the ownership concept remove the need for a garbage collector in the runtime environment, it does not require the programmer to explicitly do certain safety tasks~\cite{qin2020understanding}. Due to the memory and thread safety guarantees that Rust provides, many software companies adopted Rust in production. For example, Dropbox decided to migrate to Rust for developing their storage system, since they were not completely satisfied with the features provided by AWS’s S3 or Golang~\cite{dropbox}.

### 2022-10-16 sections/methodology.tex
added abstract/conclusion; fixed results a bit

The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and encode the important parts of the derived AST in fixed-size datapoints. Work in this vein makes the assumption that each cluster corresponds to one change pattern \cite{hanam2016discovering} \cite{campos2019discovering} \cite{yang2022mining}.

@@

Our target repositories were the top 18 most-starred open source Rust projects on GitHub. We mined (using Pydriller \cite{spadini2018pydriller}) all the bug related commits and ran them through our pipeline. Then, we applied the DBSCAN clustering algorithm \cite{ester1996density} on the resulting datapoints. In Section~\ref{sec:clustering_data} we justify our choice of DBSCAN for clustering, and how we tuned it to improve cluster quality.

@@

\paragraph{Encoding the data.} A key question is how to transform these paths to a fixed size datapoint. First, we need to define a fixed set of columns. We need fixed-size datapoints, as they are a requirement for both centroid-based and density-based well-known clustering algorithms \cite{xu2005survey}. The number of columns would depend on the amount of information we want to encode in the datapoints.

@@

A reasonable work around would be to reduce $n$ from the full set of non terminals to a smaller set of categories (e.g. a category for larger entites like class or function definitions, and a category for smaller entities like statements and expressions), and then to record all possible combinations of these categories (similar to \cite{hanam2016discovering}). That approach reduces the number of dimensions and provides more reasonable and syntactically-correct combinations. However, it would still yield a sparse dataset.

@@

We used the Pydriller library to mine software repositories. Pydriller provides APIs to extract commits from a Git repository and to search through different revisions of files. Algorithm~\ref{alg} shows how we used Pydriller to run the target repositories through our code analysis pipeline and populate two databases: one for borrow-checker related patterns (which involve items in \texttt{BCE}) and one for general patterns (which don't). For each repository, we search all commits that include bug-fixing related keywords within their commit messages. Here, we chose bug-fixing related keywords based on the set introduced in \cite{zhang2018empirical}, excluding `nan' and `inf'.

@@

Having collected our datapoints, our next task is to cluster them so that we can categorize bug patterns. Like~\cite{hanam2016discovering}, we use the DBSCAN clustering algorithm for two main reasons. Firstly, in contrast to k-means, it does not require the number of clusters in advance as an input to the algorithm. Secondly, it is a density-based clustering method, which means that it detects arbitrarily shaped clusters; centroid-based methods like k-means can't detect such clusters. We tried k-means and the more powerful fuzzy variant Fuzzy C-means \cite{dunn1973fuzzy} and did not get useful clusters.

@@

Similar to \cite{yang2022mining}, following Cotroneo et al. \cite{cotroneo2019analyzing} definitions, we divide the clusters in three different groups:

@@

In our manual analysis, we randomly picked 50 datapoints of each cluster. If the cluster had fewer than 50 datapoints, we would analyze all of the datapoints in the cluster. Next, we read the code and the bug reports related to all the datapoints (if they existed). If we judged that the members of a cluster fit into a pattern, we wrote a natural language description of that cluster. Considering the descriptions, we linked similar clusters together as possible candidates for merging. We considered these candidates and carried out appropriate ones. Finally, we selected 19 of the remaining clusters---the ones that manifested concrete changes. We discuss about our findings in Sections~\ref{sec:common_patterns} and \ref{sec:bc_patterns}.

>>>>

The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and encode the important parts of the derived AST in fixed-size datapoints. Work in this vein makes the assumption that each cluster corresponds to one change pattern~\cite{hanam2016discovering,campos2019discovering,yang2022mining}.

@@

Our target repositories were the top 18 most-starred open source Rust projects on GitHub. We mined (using Pydriller~\cite{spadini2018pydriller}) all the bug related commits and ran them through our pipeline. Then, we applied the DBSCAN clustering algorithm~\cite{ester1996density} on the resulting datapoints. In Section~\ref{sec:clustering_data} we justify our choice of DBSCAN for clustering, and how we tuned it to improve cluster quality.

@@

\paragraph{Encoding the data.} A key question is how to transform these paths to a fixed size datapoint. First, we need to define a fixed set of columns. We need fixed-size datapoints, as they are a requirement for both centroid-based and density-based well-known clustering algorithms~\cite{xu2005survey}. The number of columns would depend on the amount of information we want to encode in the datapoints.

@@

A reasonable work around would be to reduce $n$ from the full set of non terminals to a smaller set of categories (e.g. a category for larger entites like class or function definitions, and a category for smaller entities like statements and expressions), and then to record all possible combinations of these categories (similar to~\cite{hanam2016discovering}). That approach reduces the number of dimensions and provides more reasonable and syntactically-correct combinations. However, it would still yield a sparse dataset.

@@

We used the Pydriller library to mine software repositories. Pydriller provides APIs to extract commits from a Git repository and to search through different revisions of files. Algorithm~\ref{alg} shows how we used Pydriller to run the target repositories through our code analysis pipeline and populate two databases: one for borrow-checker related patterns (which involve items in \texttt{BCE}) and one for general patterns (which don't). For each repository, we search all commits that include bug-fixing related keywords within their commit messages. Here, we chose bug-fixing related keywords based on the set introduced in~\cite{zhang2018empirical}, excluding `nan' and `inf'.

@@

Having collected our datapoints, our next task is to cluster them so that we can categorize bug patterns. Like~\cite{hanam2016discovering}, we use the DBSCAN clustering algorithm for two main reasons. Firstly, in contrast to k-means, it does not require the number of clusters in advance as an input to the algorithm. Secondly, it is a density-based clustering method, which means that it detects arbitrarily shaped clusters; centroid-based methods like k-means can't detect such clusters. We tried k-means and the more powerful fuzzy variant Fuzzy C-means~\cite{dunn1973fuzzy} and did not get useful clusters.

@@

Similar to~\cite{yang2022mining}, following Cotroneo et al.~\cite{cotroneo2019analyzing} definitions, we divide the clusters in three different groups:

@@

In our manual analysis, we randomly picked 50 datapoints of each cluster. If the cluster had fewer than 50 datapoints, we would analyze all of the datapoints in the cluster. Next, we read the code and the bug reports related to all the datapoints (if they existed). If we judged that the members of a cluster fit into a pattern, we wrote a natural language description of that cluster. Considering the descriptions, we linked similar clusters together as possible candidates for merging. We considered these candidates and carried out appropriate ones. Finally, we selected 19 of the remaining clusters---the ones that manifested concrete changes. We discuss about our findings in Sections~\ref{sec:common_patterns}~and~\ref{sec:bc_patterns}.

### 2022-10-16 sections/results.tex
added abstract/conclusion; fixed results a bit

The general idea for our evaluation is whether the most important program elements get the high values in our encoding. The analyzer compares the actual change in the code and compares it to the visual representation of its respective data point. We used circle pack figures to visualize each data point \cite{collins2003circle}. For this purpose, we used the packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}. In these figures, the radius of the circle of each non terminal shows its importance in our encoding (similar to Figure~\ref{fig:essence}).

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
// G.attr.struct.add: Adding attributes to a struct
++ #[derive(Debug, Clone)]
pub struct Call {
/// body 
}

//G.attr.struct.drop: Removing attributes from a struct
-- #[derive(Debug, Clone)]
pub struct Call {
/// body 
}

// G.attr.struct.change: Changing an existing attribute of a struct
-- #[derive(Debug)]
++ #[derive(Debug, Clone)]
pub struct Call {
/// body 
}
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
// G.attr.struct.field.add: Adding attributes to struct fields
#[derive(Serialize)]
pub struct Retain {
    ++ #[serde(skip_serializing_if = "is_empty")]
    pub attributes: Option<Attributes>,
}

// G.attr.struct.field.drop: Removing attributes from struct fields
#[derive(Serialize)]
pub struct Retain {
    -- #[serde(skip_serializing_if = "Option::is_none")]
    pub attributes: Option<Attributes>,
}

// G.attr.struct.field.change: Changing attributes of struct fields
#[derive(Serialize)]
pub struct Retain {
    -- #[serde(skip_serializing_if = "Option::is_none")]
    ++ #[serde(skip_serializing_if = "is_empty")]
    pub attributes: Option<Attributes>,
}
\end{lstlisting}

@@

// G.struct.field.pub.add: Adding pub to a field
pub struct HtmlBlock {
    -- content: BlockContent,
    ++ pub content: BlockContent,
    brace: token::Brace,
}

// G.struct.field.pub.drop: Removing pub from a field
pub struct Languages {
--  pub named: HashMap<LanguageId, Arc<LanguageDefinition>>,
++  named: HashMap<LanguageId, Arc<LanguageDefinition>>,
    extensions: HashMap<String, Arc<LanguageDefinition>>,
}

@@

#[ast_node("MediaRule")]
pub struct MediaRule {
    pub span: Span,
    -- pub media: MediaQueryList,
    ++ pub media: Option<MediaQueryList>,
    pub rules: Vec<Rule>,
}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
pub struct Manifest {
  -- pub features: Vec<String>,
  ++ pub features: HashSet<String>,
}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
#[derive(Debug, Clone, PartialEq)]
pub struct Anchor {
    -- point: Point<usize>,
    ++ point: Point<isize>,
    side: Side,
}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
#[derive(Debug, Clone, PartialEq, Eq)]
pub enum ResolvedDependency {
  Resolved(ModuleSpecifier),
  -- Err(String),
  ++ Err(ResolvedDependencyErr),
}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
#[allow(dead_code)]
pub struct CoreState {
    // Other fields
    -- pending_views: Vec<ViewId>,
    ++ pending_views: Vec<(ViewId, Table)>,
    peer: Client
}
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
-- pub struct EventLoop<T: tty::EventedReadWrite> {
++ pub struct EventLoop<T: tty::EventedPty> {
    poll: mio::Poll,
    // Other fields
}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
start_plugin_process(
--  manifest.clone(),
++  manifest,
    self.next_plugin_id(),
    self.self_ref.as_ref().unwrap().clone(),
);
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
-- let field_id = schema.get_or_create(attribute.clone())?;
++ let field_id = schema.get_or_create(&attribute)?;
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
let repo = replace(&mut *contents, Processing).inner_repo();
-- let statuses = repo_to_statuses(repo, &self.workdir);
++ let statuses = repo_to_statuses(&repo, &self.workdir);
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
enums.iter().for_each(|e| {
    let mut enum_template = EnumTemplate::new();
    -- enum_template.set_message_enum(&e);
    ++ enum_template.set_message_enum(e);
    // Other statements
});
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
for p in wixobjs {
--  args.push(p.to_string());
++  args.push((*p).to_string());
}

let displayed_attributes = schema.clone().map(|s| {
    s.displayed_name()
        .iter()
--      .map(|s| (*s).to_string())
++      .map(|s| s.to_string())
        .collect::<HashSet<String>>()
});
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
parser::Parser::new(args)
        .parse_module()
--      .map_err(|e| {
++      .map_err(|mut e| {
            e.emit();
            ()
        })?
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
-- let mut tx = tx.clone();
++ let tx = tx.clone();
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
pub fn expand_delimited_square(
--  children: &Vec<TokenNode>,
++  children: &[TokenNode],
) -> Result<hir::Expression, ParseError> {
    // body
}
\end{lstlisting}
\noindent\textbf{Description:} In struct field definition or the types of the function formal parameters, type \verb+&Vec<T>+ changes to \verb+&[T]+. Clippy can detect the opportunity to apply this pattern. The reason is reducing type complication, as slice types in the form of \verb+&[T]+ or \verb+&str+ are usually sufficient for most use cases.

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
-- const QUEUE_SIZES: &'static [u16] = &[QUEUE_SIZE];
++ const QUEUE_SIZES: &[u16] = &[QUEUE_SIZE];

>>>>

The general idea for our evaluation is whether the most important program elements get the high values in our encoding. The analyzer compares the actual change in the code and compares it to the visual representation of its respective data point. We used circle pack figures to visualize each data point~\cite{collins2003circle}. For this purpose, we used the packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}. In these figures, the radius of the circle of each non terminal shows its importance in our encoding (similar to Figure~\ref{fig:essence}).

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
// G.attr.struct.add: Adding attributes to a struct
<@\textcolor{green}{++}@> #[derive(Debug, Clone)]
pub struct Call {
/// body 
}

//G.attr.struct.drop: Removing attributes from a struct
<@\textcolor{red}{--}@> #[derive(Debug, Clone)]
pub struct Call {
/// body 
}

// G.attr.struct.change: Changing an existing attribute of a struct
<@\textcolor{red}{--}@> #[derive(Debug)]
<@\textcolor{green}{++}@> #[derive(Debug, Clone)]
pub struct Call {
/// body 
}
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
// G.attr.struct.field.add: Adding attributes to struct fields
#[derive(Serialize)]
pub struct Retain {
    <@\textcolor{green}{++}@> #[serde(skip_serializing_if = "is_empty")]
    pub attributes: Option<Attributes>,
}

// G.attr.struct.field.drop: Removing attributes from struct fields
#[derive(Serialize)]
pub struct Retain {
    <@\textcolor{red}{--}@> #[serde(skip_serializing_if = "Option::is_none")]
    pub attributes: Option<Attributes>,
}

// G.attr.struct.field.change: Changing attributes of struct fields
#[derive(Serialize)]
pub struct Retain {
    <@\textcolor{red}{--}@> #[serde(skip_serializing_if = "Option::is_none")]
    <@\textcolor{green}{++}@> #[serde(skip_serializing_if = "is_empty")]
    pub attributes: Option<Attributes>,
}
\end{lstlisting}

@@

// G.struct.field.pub.add: Adding pub to a field
pub struct HtmlBlock {
    <@\textcolor{red}{--}@> content: BlockContent,
    <@\textcolor{green}{++}@> pub content: BlockContent,
    brace: token::Brace,
}

// G.struct.field.pub.drop: Removing pub from a field
pub struct Languages {
<@\textcolor{red}{--}@>  pub named: HashMap<LanguageId, Arc<LanguageDefinition>>,
<@\textcolor{green}{++}@>  named: HashMap<LanguageId, Arc<LanguageDefinition>>,
    extensions: HashMap<String, Arc<LanguageDefinition>>,
}

@@

#[ast_node("MediaRule")]
pub struct MediaRule {
    pub span: Span,
    <@\textcolor{red}{--}@> pub media: MediaQueryList,
    <@\textcolor{green}{++}@> pub media: Option<MediaQueryList>,
    pub rules: Vec<Rule>,
}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
pub struct Manifest {
  <@\textcolor{red}{--}@> pub features: Vec<String>,
  <@\textcolor{green}{++}@> pub features: HashSet<String>,
}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
#[derive(Debug, Clone, PartialEq)]
pub struct Anchor {
    <@\textcolor{red}{--}@> point: Point<usize>,
    <@\textcolor{green}{++}@> point: Point<isize>,
    side: Side,
}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
#[derive(Debug, Clone, PartialEq, Eq)]
pub enum ResolvedDependency {
  Resolved(ModuleSpecifier),
  <@\textcolor{red}{--}@> Err(String),
  <@\textcolor{green}{++}@> Err(ResolvedDependencyErr),
}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
#[allow(dead_code)]
pub struct CoreState {
    // Other fields
    <@\textcolor{red}{--}@> pending_views: Vec<ViewId>,
    <@\textcolor{green}{++}@> pending_views: Vec<(ViewId, Table)>,
    peer: Client
}
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
<@\textcolor{red}{--}@> pub struct EventLoop<T: tty::EventedReadWrite> {
<@\textcolor{green}{++}@> pub struct EventLoop<T: tty::EventedPty> {
    poll: mio::Poll,
    // Other fields
}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
start_plugin_process(
<@\textcolor{red}{--}@>  manifest.clone(),
<@\textcolor{green}{++}@>  manifest,
    self.next_plugin_id(),
    self.self_ref.as_ref().unwrap().clone(),
);
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
<@\textcolor{red}{--}@> let field_id = schema.get_or_create(attribute.clone())?;
<@\textcolor{green}{++}@> let field_id = schema.get_or_create(&attribute)?;
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
let repo = replace(&mut *contents, Processing).inner_repo();
<@\textcolor{red}{--}@> let statuses = repo_to_statuses(repo, &self.workdir);
<@\textcolor{green}{++}@> let statuses = repo_to_statuses(&repo, &self.workdir);
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
enums.iter().for_each(|e| {
    let mut enum_template = EnumTemplate::new();
    <@\textcolor{red}{--}@> enum_template.set_message_enum(&e);
    <@\textcolor{green}{++}@> enum_template.set_message_enum(e);
    // Other statements
});
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
for p in wixobjs {
<@\textcolor{red}{--}@>  args.push(p.to_string());
<@\textcolor{green}{++}@>  args.push((*p).to_string());
}

let displayed_attributes = schema.clone().map(|s| {
    s.displayed_name()
        .iter()
<@\textcolor{red}{--}@>      .map(|s| (*s).to_string())
<@\textcolor{green}{++}@>      .map(|s| s.to_string())
        .collect::<HashSet<String>>()
});
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
parser::Parser::new(args)
        .parse_module()
<@\textcolor{red}{--}@>      .map_err(|e| {
<@\textcolor{green}{++}@>      .map_err(|mut e| {
            e.emit();
            ()
        })?
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
<@\textcolor{red}{--}@> let mut tx = tx.clone();
<@\textcolor{green}{++}@> let tx = tx.clone();
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
pub fn expand_delimited_square(
<@\textcolor{red}{--}@>  children: &Vec<TokenNode>,
<@\textcolor{green}{++}@>  children: &[TokenNode],
) -> Result<hir::Expression, ParseError> {
    // body
}
\end{lstlisting}
\noindent\textbf{Description:} In struct field definition or the types of the function formal parameters, type \verb+&Vec<T>+ changes to \verb+&[T]+. Clippy can detect the opportunity to apply this pattern. The reason is reducing type complication, as slice types in the form of \verb+&[T]+ or \verb+&str+ are usually sufficient for most use cases.

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
<@\textcolor{red}{--}@> const QUEUE_SIZES: &'static [u16] = &[QUEUE_SIZE];
<@\textcolor{green}{++}@> const QUEUE_SIZES: &[u16] = &[QUEUE_SIZE];
